---
title: "Mawi vs. Frozen Stool Data"
author: "Rigel Sison"
date: "1/4/2022"
output: html_document
---

```{r}
# Load packages
necessary_packages <-  c("directlabels", "knitr", "clustsig", "ellipse", "phyloseq", "ggplot2", "plyr", "vegan", "reshape2", "ampvis2", "DESeq2", "gtable", "gridExtra", "grid", "Biostrings", "data.table", "labdsv", "ape", "multcompView")
packages <- lapply(necessary_packages, library, character.only = TRUE)
#("biomformat)
library(dplyr)
library(decontam)
library("RColorBrewer")
library("magrittr")
library("scales")
library("randomForest")
library(ggtern)
library(tidyr)
library(viridis)
library(cowplot)
library(tidyverse) # Easily Install and Load the 'Tidyverse'.
library(coin) # Conditional Inference Procedures in a Permutation Test Framework.
library(reshape2) # Flexibly Reshape Data: A Reboot of the Reshape Package.
library(ggnewscale)
library(VennDiagram)
library(gplots)
library(tibble)
```

```{r}
table <- read.table(file = 'frozen_v_mawi.tsv', sep = '\t', header = TRUE)
copy <- table
rownames(copy) <- c(1:733)
rownames(copy)[1:9] <- paste0("00", rownames(copy)[1:9])
rownames(copy)[10:99] <- paste0("0", rownames(copy)[10:99])
rownames(copy) <- paste0("ST_", rownames(copy))
colnames(copy) <- gsub('MP_','', colnames(copy))
```

```{r}
final <- copy
st.phylo <- final[, 2:8]
st.counts <- final[, 9:74]
st.counts <- trunc(st.counts)
```

```{r}
st.counts.t <- t(st.counts)

map <- read.table("map_frozen_mawi.txt", header=TRUE, fill = TRUE)
map$Date <- as.POSIXct(map$Date,
                       format = "%Y-%m-%d")
rownames(map) <- map[,1]
map.samp <- map[,2:ncol(map)]
map.samp <- map.samp[rownames(st.counts.t),]

identical(rownames(st.counts.t),rownames(map))

st.phylo <- as.matrix(st.phylo)

class(st.counts.t) <- "numeric"

OTU <- otu_table(st.counts, taxa_are_rows = TRUE)
TAX <- tax_table(st.phylo)
META <- sample_data(map)
st.physeq <- phyloseq(OTU, TAX, META)

```

```{r}
#ex1 = prune_taxa(row.names.remove, st.physeq)
```

```{r}

sum(taxa_sums(st.physeq) == 0)
sum(taxa_sums(st.physeq) == 1) #7
sum(taxa_sums(st.physeq) == 2) #13
sum(taxa_sums(st.physeq) == 3) #11

st.physeq.f <- filter_taxa(st.physeq, function(x) sum(x > 3) > (0.02*length(x)), TRUE)

```

Subset the phyloseq data by RNA and DNA

```{r}

st.physeq.dna <- subset_samples(st.physeq.f, Nucleic.Acid == "DNA")
st.physeq.rna <- subset_samples(st.physeq.f, Nucleic.Acid == "RNA")

```

```{r}
library(MicrobiotaProcess)
vennlist <- get_vennlist(obj=st.physeq.dna, factorNames="Type")
ItemsList<- venn(vennlist, show.plot = FALSE)
intersections<- attributes(ItemsList)$intersections
detach("package:MicrobiotaProcess", unload = TRUE)
```

```{r}
#filter OTUs not present in both day 0 samples across Mawi and Frozen
row.names.remove <- c(intersections$Mawi, intersections$Frozen)
copy.f <- copy[!(rownames(copy) %in% row.names.remove), ]
```

```{r}
final <- copy.f
st.phylo <- final[, 2:8]
st.counts <- final[, 9:74]
st.counts <- trunc(st.counts)
```

```{r}
st.counts.t <- t(st.counts)
map <- read.table("map_frozen_mawi.txt", header=TRUE, fill = TRUE)
rownames(map) <- map[,1]
map.samp <- map[,2:ncol(map)]
map.samp <- map.samp[rownames(st.counts.t),]

identical(rownames(st.counts.t),rownames(map))

st.phylo <- as.matrix(st.phylo)

class(st.counts.t) <- "numeric"

OTU <- otu_table(st.counts, taxa_are_rows = TRUE)
TAX <- tax_table(st.phylo)
META <- sample_data(map)
st.physeq <- phyloseq(OTU, TAX, META)

```

```{r}

sum(taxa_sums(st.physeq) == 0)
sum(taxa_sums(st.physeq) == 1) #7
sum(taxa_sums(st.physeq) == 2) #13
sum(taxa_sums(st.physeq) == 3) #11

st.physeq.f <- filter_taxa(st.physeq, function(x) sum(x > 3) > (0.02*length(x)), TRUE)

```

Subset the phyloseq data by RNA and DNA

```{r}

st.physeq.dna <- subset_samples(st.physeq.f, Nucleic.Acid == "DNA")
st.physeq.rna <- subset_samples(st.physeq.f, Nucleic.Acid == "RNA")

```

```{r}

sample_data(st.physeq.f)$Nucleic.Acid <- factor(sample_data(st.physeq.f)$Nucleic.Acid, levels = c('RNA', 'DNA'))

df <- data.frame(sample_data(st.physeq.f))
df$LibrarySize <- sample_sums(st.physeq.f)
df <- df[order(df$LibrarySize),]
df$ID <- seq(nrow(df))

p_ls_nucleic.acid <- ggplot(data = df, aes(x=ID, y=LibrarySize, color=Nucleic.Acid)) + geom_point() +
  scale_color_manual(values = c('RNA' = 'black', 'DNA' = 'red'))

```



Venn Diagrams
```{r}

st.physeq.dna.day_0 <- subset_samples(st.physeq.dna, Day == "0")
st.physeq.dna.day_3 <- subset_samples(st.physeq.dna, Day == "3")
st.physeq.dna.day_7 <- subset_samples(st.physeq.dna, Day == "7")
st.physeq.dna.day_14 <- subset_samples(st.physeq.dna, Day == "14")
st.physeq.dna.day_21 <- subset_samples(st.physeq.dna, Day == "21")
st.physeq.dna.day_40 <- subset_samples(st.physeq.dna, Day == "40")
library(MicrobiotaProcess)
vennlist <- get_vennlist(obj=st.physeq.dna.day_0, factorNames="Type")
vennp.day_0 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("#00AED7", "#FD9347"),
                      cat.col=c("#00AED7", "#FD9347"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.day_3, factorNames="Type")
vennp.day_3 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("#00AED7", "#FD9347"),
                      cat.col=c("#00AED7", "#FD9347"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.day_7, factorNames="Type")
vennp.day_7 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("#00AED7", "#FD9347"),
                      cat.col=c("#00AED7", "#FD9347"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.day_14, factorNames="Type")
vennp.day_14 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("#00AED7", "#FD9347"),
                      cat.col=c("#00AED7", "#FD9347"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.day_21, factorNames="Type")
vennp.day_21 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("#00AED7", "#FD9347"),
                      cat.col=c("#00AED7", "#FD9347"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.day_40, factorNames="Type")
vennp.day_40 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("#00AED7", "#FD9347"),
                      cat.col=c("#00AED7", "#FD9347"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

detach("package:MicrobiotaProcess", unload = TRUE)
```

```{r}

sample_data(st.physeq.dna)$Type <- factor(sample_data(st.physeq.dna)$Type, levels = c('Frozen', 'Mawi'))

df <- 
  data.frame(sample_data(st.physeq.dna))
df$LibrarySize <- sample_sums(st.physeq.dna)
df <- df[order(df$LibrarySize),]
df$ID <- seq(nrow(df))

p_ls_dna.type <- ggplot(data = df, aes(x=ID, y=LibrarySize, color=Type)) + geom_point() +
  scale_color_manual(values = c('Frozen' = '#6BA7CC', 'Mawi' = '#ff6600'))

p_ls_dna.type
ggsave("library_size_dna.pdf",
       width = 6,
       height = 4,
       units = "in")
```

```{r}

sample_data(st.physeq.rna)$Type <- factor(sample_data(st.physeq.rna)$Type, levels = c('Frozen', 'Mawi'))

df <- 
  data.frame(sample_data(st.physeq.rna))
df$LibrarySize <- sample_sums(st.physeq.rna)
df <- df[order(df$LibrarySize),]
df$ID <- seq(nrow(df))

p_ls_rna.type <- ggplot(data = df, aes(x=ID, y=LibrarySize, color=Type)) + geom_point() +
  scale_color_manual(values = c('Frozen' = '#6BA7CC', 'Mawi' = '#ff6600'))

p_ls_rna.type
```

```{r}

sample_data(st.physeq)$Type <- factor(sample_data(st.physeq)$Type, levels = c('Frozen', 'Mawi'))

df <- data.frame(sample_data(st.physeq))
df$LibrarySize <- sample_sums(st.physeq)
df <- df[order(df$LibrarySize),]
df$ID <- seq(nrow(df))

p_ls <- ggplot(data = df, aes(x=ID, y=LibrarySize, color=Type, shape=Nucleic.Acid)) + 
  geom_point(size = 2) +
  scale_color_manual(values = c('Frozen' = '#6BA7CC', 'Mawi' = '#ff6600'))

p_ls
ggsave("library_size.pdf",
       width = 6,
       height = 4,
       units = "in")

write.csv(df, "library_size.csv", row.names = FALSE)
```

```{r}

set.seed(13)
st.physeq.fr = rarefy_even_depth(st.physeq.f, sample.size = 300000, rngseed = TRUE)

```

```{r}
#library(patchwork)
#set.seed(13)
#rareres <- get_rarecurve(obj = st.physeq.fr)
#p_rare <- ggrarecurve(obj = rareres,
#                      indexNames = c("Observe"),
#                      )
```

Bray-Curtis NMDS
```{r}

set.seed(13)
# Ordinate
st.physeq.nmds.bray <- ordinate(
  physeq = st.physeq.fr,
  method = 'NMDS',
  distance = 'bray'
)

bray_ord <- phyloseq::distance(physeq = st.physeq.fr, method = 'bray')

df_ord <- data.frame(sample_data(st.physeq.fr))
adonis(bray_ord ~ Nucleic.Acid, data = df_ord)

beta <- betadisper(bray_ord, df_ord$Nucleic.Acid)
permutest(beta)

nmds.bray <- plot_ordination(
  physeq = st.physeq.fr,
  ordination <- st.physeq.nmds.bray,
  color = 'Nucleic.Acid',
  shape = 'Type',
  title = 'nMDS of Bacterial Communities in Stool Samples (Bray-Curtis)'
) +
  scale_color_manual(values = c('#8B008B', '#00008B')) +
  geom_point(aes(color = Nucleic.Acid), size = 3) +
  theme(plot.title = element_text(hjust = 0.5)) 
```

```{r}

set.seed(13)
st.physeq.dna.r = rarefy_even_depth(st.physeq.dna, sample.size = 300000, rngseed = TRUE)

set.seed(13)
# Ordinate
st.physeq.nmds.bray.dna <- ordinate(
  physeq = st.physeq.dna.r,
  method = 'NMDS',
  distance = 'bray'
)

bray_ord <- phyloseq::distance(physeq = st.physeq.dna.r, method = 'bray')

df_ord <- data.frame(sample_data(st.physeq.dna.r))
adonis(bray_ord ~ Type, data = df_ord)
adonis(bray_ord ~ Day, data = df_ord)
adonis(bray_ord ~ Pool, data = df_ord)
adonis(bray_ord ~ Type + Day + Pool, data = df_ord)

beta <- betadisper(bray_ord, df_ord$Type)
permutest(beta)

nmds.bray.dna <- plot_ordination(
  physeq = st.physeq.dna.r,
  ordination <- st.physeq.nmds.bray.dna,
  color = 'Nucleic.Acid',
  title = 'nMDS of Bacterial Communities in Stool DNA Samples (Bray-Curtis)'
) +
  scale_color_manual(values = c('Frozen' = 'blue', 'Mawi' = '#ff6600')) +
  geom_point(aes(color = Type), size = 3) +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = 0.225, y = 0.1, label = "\n Stress=0.15\n Adonis\n R^2=0.12\n p=0.001")
  
```

```{r}

set.seed(13)
st.physeq.rna.r = rarefy_even_depth(st.physeq.rna, sample.size = 300000, rngseed = TRUE)

set.seed(13)
# Ordinate
st.physeq.nmds.bray.rna <- ordinate(
  physeq = st.physeq.rna.r,
  method = 'NMDS',
  distance = 'bray'
)

bray_ord <- phyloseq::distance(physeq = st.physeq.rna.r, method = 'bray')

df_ord <- data.frame(sample_data(st.physeq.rna.r))
adonis(bray_ord ~ Type, data = df_ord)

beta <- betadisper(bray_ord, df_ord$Type)
permutest(beta)

nmds.bray.rna <- plot_ordination(
  physeq = st.physeq.rna.r,
  ordination <- st.physeq.nmds.bray.rna,
  color = 'Nucleic.Acid',
  title = 'nMDS of Bacterial Communities in Stool RNA Samples (Bray-Curtis)'
) +
  scale_color_manual(values = c('Frozen' = 'blue', 'Mawi' = '#ff6600')) +
  geom_point(aes(color = Type), size = 3) +
  theme(plot.title = element_text(hjust = 0.5))
```
All Bray-Curtis NMDS plots

```{r}
#nmds.bray
nmds.bray.dna
ggsave("nmds_bray_dna_all.png",
       width = 6.5,
       height = 4,
       units = "in")
#nmds.bray.rna
```

Jaccard NMDS
```{r}

set.seed(13)
# Ordinate
st.physeq.nmds.jaccard <- ordinate(
  physeq = st.physeq.fr,
  method = 'NMDS',
  distance = 'jaccard'
)

jaccard_ord <- phyloseq::distance(physeq = st.physeq.fr, method = 'jaccard')

df_ord <- data.frame(sample_data(st.physeq.fr))
adonis(jaccard_ord ~ Nucleic.Acid, data = df_ord)

beta <- betadisper(jaccard_ord, df_ord$Nucleic.Acid)
permutest(beta)

nmds.jaccard <- plot_ordination(
  physeq = st.physeq.fr,
  ordination <- st.physeq.nmds.jaccard,
  color = 'Nucleic.Acid',
  shape = 'Type',
  title = 'nMDS of Bacterial Communities in Stool Samples (Jaccard)'
) +
  scale_color_manual(values = c('#8B008B', '#00008B')) +
  geom_point(aes(color = Nucleic.Acid), size = 3) +
  theme(plot.title = element_text(hjust = 0.5)) 
```

```{r}

set.seed(13)
# Ordinate
st.physeq.nmds.jaccard.dna <- ordinate(
  physeq = st.physeq.dna.r,
  method = 'NMDS',
  distance = 'jaccard'
)

jaccard_ord <- phyloseq::distance(physeq = st.physeq.dna.r, method = 'jaccard')

df_ord <- data.frame(sample_data(st.physeq.dna.r))
adonis(jaccard_ord ~ Type, data = df_ord)

beta <- betadisper(jaccard_ord, df_ord$Type)
permutest(beta)

nmds.jaccard.dna <- plot_ordination(
  physeq = st.physeq.dna.r,
  ordination <- st.physeq.nmds.jaccard.dna,
  color = 'Nucleic.Acid',
  title = 'nMDS of Bacterial Communities in Stool DNA Samples (Jaccard)'
) +
  scale_color_manual(values = c('Frozen' = 'blue', 'Mawi' = '#ff6600')) +
  geom_point(aes(color = Type), size = 3) +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = 0.3, y = 0.15, label = "\n Stress=0.15\n Adonis\n R^2=0.11\n p=0.001")
  
```

```{r}

set.seed(13)
# Ordinate
st.physeq.nmds.jaccard.rna <- ordinate(
  physeq = st.physeq.rna.r,
  method = 'NMDS',
  distance = 'jaccard'
)

jaccard_ord <- phyloseq::distance(physeq = st.physeq.rna.r, method = 'jaccard')

df_ord <- data.frame(sample_data(st.physeq.rna.r))
adonis(jaccard_ord ~ Type, data = df_ord)

beta <- betadisper(jaccard_ord, df_ord$Type)
permutest(beta)

nmds.jaccard.rna <- plot_ordination(
  physeq = st.physeq.rna.r,
  ordination <- st.physeq.nmds.jaccard.rna,
  color = 'Nucleic.Acid',
  title = 'nMDS of Bacterial Communities in Stool RNA Samples (Jaccard)'
) +
  scale_color_manual(values = c('Frozen' = 'blue', 'Mawi' = '#ff6600')) +
  geom_point(aes(color = Type), size = 3) +
  theme(plot.title = element_text(hjust = 0.5)) 
```
All Jaccard NMDS plots

```{r}
#nmds.jaccard
nmds.jaccard.dna
ggsave("nmds_jaccard_dna_all.png",
       width = 6,
       height = 4,
       units = "in")
#nmds.jaccard.rna
```

Distance Comparison across time points (Bray-Curtis)

```{r}

# T0 vs T3, Mawi, DNA, Bray

st.physeq.dna.t0_3.mawi <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '3')
dist.dna.t0_3.mawi <- phyloseq::distance(physeq = st.physeq.dna.t0_3.mawi, method = 'bray')
dist.dna.t0_3.mawi.df <- 
  matrix(dist.dna.t0_3.mawi)
dist.dna.t0_3.mawi.list <- reshape2::melt(dist.dna.t0_3.mawi.df)[reshape2::melt(upper.tri(dist.dna.t0_3.mawi.df))$value,]
names(dist.dna.t0_3.mawi.list) <- c("c1", "c2", "distance")
dist.dna.t0_3.mawi.list <- cbind(dist.dna.t0_3.mawi.list, "comp" = "T0_vs_T03", "type" = "Mawi")

# T0 vs T7, Mawi, DNA, Bray

st.physeq.dna.t0_7.mawi <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '7')
dist.dna.t0_7.mawi <- phyloseq::distance(physeq = st.physeq.dna.t0_7.mawi, method = 'bray')
dist.dna.t0_7.mawi.df <- 
  matrix(dist.dna.t0_7.mawi)
dist.dna.t0_7.mawi.list <- reshape2::melt(dist.dna.t0_7.mawi.df)[reshape2::melt(upper.tri(dist.dna.t0_7.mawi.df))$value,]
names(dist.dna.t0_7.mawi.list) <- c("c1", "c2", "distance")
dist.dna.t0_7.mawi.list <- cbind(dist.dna.t0_7.mawi.list, "comp" = "T0_vs_T07", "type" = "Mawi")

# T0 vs T14, Mawi, DNA, Bray

st.physeq.dna.t0_14.mawi <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '14')
dist.dna.t0_14.mawi <- phyloseq::distance(physeq = st.physeq.dna.t0_14.mawi, method = 'bray')
dist.dna.t0_14.mawi.df <- 
  matrix(dist.dna.t0_14.mawi)
dist.dna.t0_14.mawi.list <- reshape2::melt(dist.dna.t0_14.mawi.df)[reshape2::melt(upper.tri(dist.dna.t0_14.mawi.df))$value,]
names(dist.dna.t0_14.mawi.list) <- c("c1", "c2", "distance")
dist.dna.t0_14.mawi.list <- cbind(dist.dna.t0_14.mawi.list, "comp" = "T0_vs_T14", "type" = "Mawi")

# T0 vs T21, Mawi, DNA, Bray

st.physeq.dna.t0_21.mawi <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '21')
dist.dna.t0_21.mawi <- phyloseq::distance(physeq = st.physeq.dna.t0_21.mawi, method = 'bray')
dist.dna.t0_21.mawi.df <- 
  matrix(dist.dna.t0_21.mawi)
dist.dna.t0_21.mawi.list <- reshape2::melt(dist.dna.t0_21.mawi.df)[reshape2::melt(upper.tri(dist.dna.t0_21.mawi.df))$value,]
names(dist.dna.t0_21.mawi.list) <- c("c1", "c2", "distance")
dist.dna.t0_21.mawi.list <- cbind(dist.dna.t0_21.mawi.list, "comp" = "T0_vs_T21", "type" = "Mawi")

# T0 vs T40, Mawi, DNA, Bray

st.physeq.dna.t0_40.mawi <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '40')
dist.dna.t0_40.mawi <- phyloseq::distance(physeq = st.physeq.dna.t0_40.mawi, method = 'bray')
dist.dna.t0_40.mawi.df <- 
  matrix(dist.dna.t0_40.mawi)
dist.dna.t0_40.mawi.list <- reshape2::melt(dist.dna.t0_40.mawi.df)[reshape2::melt(upper.tri(dist.dna.t0_40.mawi.df))$value,]
names(dist.dna.t0_40.mawi.list) <- c("c1", "c2", "distance")
dist.dna.t0_40.mawi.list <- cbind(dist.dna.t0_40.mawi.list, "comp" = "T0_vs_T40", "type" = "Mawi")

#combine all tables
dist.per.time_comp.dna.mawi <- rbind(dist.dna.t0_3.mawi.list, dist.dna.t0_7.mawi.list, dist.dna.t0_14.mawi.list, dist.dna.t0_21.mawi.list, dist.dna.t0_40.mawi.list)

#plot
boxplot.dna.mawi <- ggplot(dist.per.time_comp.dna.mawi, aes(x = comp, y = distance, fill = comp)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") + theme_bw() + theme(axis.title.x=element_blank(), legend.position = "none") +
  coord_cartesian(ylim = c(0, 0.27)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

Frozen DNA

```{r}

# T0 vs T3, Frozen, DNA, Bray

st.physeq.dna.t0_3.frozen <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '3')
dist.dna.t0_3.frozen <- phyloseq::distance(physeq = st.physeq.dna.t0_3.frozen, method = 'bray')
dist.dna.t0_3.frozen.df <- 
  matrix(dist.dna.t0_3.frozen)
dist.dna.t0_3.frozen.list <- reshape2::melt(dist.dna.t0_3.frozen.df)[reshape2::melt(upper.tri(dist.dna.t0_3.frozen.df))$value,]
names(dist.dna.t0_3.frozen.list) <- c("c1", "c2", "distance")
dist.dna.t0_3.frozen.list <- cbind(dist.dna.t0_3.frozen.list, "comp" = "T0_vs_T03", "type" = "Frozen")

# T0 vs T7, Frozen, DNA, Bray

st.physeq.dna.t0_7.frozen <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '7')
dist.dna.t0_7.frozen <- phyloseq::distance(physeq = st.physeq.dna.t0_7.frozen, method = 'bray')
dist.dna.t0_7.frozen.df <- 
  matrix(dist.dna.t0_7.frozen)
dist.dna.t0_7.frozen.list <- reshape2::melt(dist.dna.t0_7.frozen.df)[reshape2::melt(upper.tri(dist.dna.t0_7.frozen.df))$value,]
names(dist.dna.t0_7.frozen.list) <- c("c1", "c2", "distance")
dist.dna.t0_7.frozen.list <- cbind(dist.dna.t0_7.frozen.list, "comp" = "T0_vs_T07", "type" = "Frozen")

# T0 vs T14, Frozen, DNA, Bray

st.physeq.dna.t0_14.frozen <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '14')
dist.dna.t0_14.frozen <- phyloseq::distance(physeq = st.physeq.dna.t0_14.frozen, method = 'bray')
dist.dna.t0_14.frozen.df <- 
  matrix(dist.dna.t0_14.frozen)
dist.dna.t0_14.frozen.list <- reshape2::melt(dist.dna.t0_14.frozen.df)[reshape2::melt(upper.tri(dist.dna.t0_14.frozen.df))$value,]
names(dist.dna.t0_14.frozen.list) <- c("c1", "c2", "distance")
dist.dna.t0_14.frozen.list <- cbind(dist.dna.t0_14.frozen.list, "comp" = "T0_vs_T14", "type" = "Frozen")

# T0 vs T21, Frozen, DNA, Bray

st.physeq.dna.t0_21.frozen <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '21')
dist.dna.t0_21.frozen <- phyloseq::distance(physeq = st.physeq.dna.t0_21.frozen, method = 'bray')
dist.dna.t0_21.frozen.df <- 
  matrix(dist.dna.t0_21.frozen)
dist.dna.t0_21.frozen.list <- reshape2::melt(dist.dna.t0_21.frozen.df)[reshape2::melt(upper.tri(dist.dna.t0_21.frozen.df))$value,]
names(dist.dna.t0_21.frozen.list) <- c("c1", "c2", "distance")
dist.dna.t0_21.frozen.list <- cbind(dist.dna.t0_21.frozen.list, "comp" = "T0_vs_T21", "type" = "Frozen")

# T0 vs T40, Frozen, DNA, Bray

st.physeq.dna.t0_40.frozen <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '40')
dist.dna.t0_40.frozen <- phyloseq::distance(physeq = st.physeq.dna.t0_40.frozen, method = 'bray')
dist.dna.t0_40.frozen.df <- 
  matrix(dist.dna.t0_40.frozen)
dist.dna.t0_40.frozen.list <- reshape2::melt(dist.dna.t0_40.frozen.df)[reshape2::melt(upper.tri(dist.dna.t0_40.frozen.df))$value,]
names(dist.dna.t0_40.frozen.list) <- c("c1", "c2", "distance")
dist.dna.t0_40.frozen.list <- cbind(dist.dna.t0_40.frozen.list, "comp" = "T0_vs_T40", "type" = "Frozen")

#combine all tables
dist.per.time_comp.dna.frozen <- rbind(dist.dna.t0_3.frozen.list, dist.dna.t0_7.frozen.list, dist.dna.t0_14.frozen.list, dist.dna.t0_21.frozen.list, dist.dna.t0_40.frozen.list)

#plot
boxplot.dna.frozen <- ggplot(dist.per.time_comp.dna.frozen, aes(x = comp, y = distance, fill = comp)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") + theme_bw() + theme(axis.title.x=element_blank(), legend.position = "none") +
  coord_cartesian(ylim = c(0, 0.27)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
  
```

```{r}
dist.per.time_comp.dna <- rbind(dist.per.time_comp.dna.frozen, dist.per.time_comp.dna.mawi)

boxplot.dna.all <- ggplot(dist.per.time_comp.dna, aes(x = comp, y = distance, fill = comp)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") +
  facet_grid(. ~ type) +
  theme_bw() +
  theme(axis.title.x = element_blank(), legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  ggtitle("Distance Across Time Points (Bray-Curtis)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
boxplot.dna.all
ggsave("boxplot_across_time_points (bray) .png",
       width = 6,
       height = 4,
       units = "in")
```

Distance Comparison across time points (Jaccard)

```{r}

# T0 vs T3, Mawi, DNA, Jaccard

st.physeq.dna.t0_3.mawi <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '3')
dist.dna.t0_3.mawi <- phyloseq::distance(physeq = st.physeq.dna.t0_3.mawi, method = 'jaccard')
dist.dna.t0_3.mawi.df <- 
  matrix(dist.dna.t0_3.mawi)
dist.dna.t0_3.mawi.list <- reshape2::melt(dist.dna.t0_3.mawi.df)[reshape2::melt(upper.tri(dist.dna.t0_3.mawi.df))$value,]
names(dist.dna.t0_3.mawi.list) <- c("c1", "c2", "distance")
dist.dna.t0_3.mawi.list <- cbind(dist.dna.t0_3.mawi.list, "comp" = "T0_vs_T03", "type" = "Mawi")

# T0 vs T7, Mawi, DNA, Jaccard

st.physeq.dna.t0_7.mawi <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '7')
dist.dna.t0_7.mawi <- phyloseq::distance(physeq = st.physeq.dna.t0_7.mawi, method = 'jaccard')
dist.dna.t0_7.mawi.df <- 
  matrix(dist.dna.t0_7.mawi)
dist.dna.t0_7.mawi.list <- reshape2::melt(dist.dna.t0_7.mawi.df)[reshape2::melt(upper.tri(dist.dna.t0_7.mawi.df))$value,]
names(dist.dna.t0_7.mawi.list) <- c("c1", "c2", "distance")
dist.dna.t0_7.mawi.list <- cbind(dist.dna.t0_7.mawi.list, "comp" = "T0_vs_T07", "type" = "Mawi")

# T0 vs T14, Mawi, DNA, Jaccard

st.physeq.dna.t0_14.mawi <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '14')
dist.dna.t0_14.mawi <- phyloseq::distance(physeq = st.physeq.dna.t0_14.mawi, method = 'jaccard')
dist.dna.t0_14.mawi.df <- 
  matrix(dist.dna.t0_14.mawi)
dist.dna.t0_14.mawi.list <- reshape2::melt(dist.dna.t0_14.mawi.df)[reshape2::melt(upper.tri(dist.dna.t0_14.mawi.df))$value,]
names(dist.dna.t0_14.mawi.list) <- c("c1", "c2", "distance")
dist.dna.t0_14.mawi.list <- cbind(dist.dna.t0_14.mawi.list, "comp" = "T0_vs_T14", "type" = "Mawi")

# T0 vs T21, Mawi, DNA, Jaccard

st.physeq.dna.t0_21.mawi <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '21')
dist.dna.t0_21.mawi <- phyloseq::distance(physeq = st.physeq.dna.t0_21.mawi, method = 'jaccard')
dist.dna.t0_21.mawi.df <- 
  matrix(dist.dna.t0_21.mawi)
dist.dna.t0_21.mawi.list <- reshape2::melt(dist.dna.t0_21.mawi.df)[reshape2::melt(upper.tri(dist.dna.t0_21.mawi.df))$value,]
names(dist.dna.t0_21.mawi.list) <- c("c1", "c2", "distance")
dist.dna.t0_21.mawi.list <- cbind(dist.dna.t0_21.mawi.list, "comp" = "T0_vs_T21", "type" = "Mawi")

# T0 vs T40, Mawi, DNA, Jaccard

st.physeq.dna.t0_40.mawi <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '40')
dist.dna.t0_40.mawi <- phyloseq::distance(physeq = st.physeq.dna.t0_40.mawi, method = 'jaccard')
dist.dna.t0_40.mawi.df <- 
  matrix(dist.dna.t0_40.mawi)
dist.dna.t0_40.mawi.list <- reshape2::melt(dist.dna.t0_40.mawi.df)[reshape2::melt(upper.tri(dist.dna.t0_40.mawi.df))$value,]
names(dist.dna.t0_40.mawi.list) <- c("c1", "c2", "distance")
dist.dna.t0_40.mawi.list <- cbind(dist.dna.t0_40.mawi.list, "comp" = "T0_vs_T40", "type" = "Mawi")

#combine all tables
dist.per.time_comp.dna.mawi <- rbind(dist.dna.t0_3.mawi.list, dist.dna.t0_7.mawi.list, dist.dna.t0_14.mawi.list, dist.dna.t0_21.mawi.list, dist.dna.t0_40.mawi.list)

#plot
boxplot.dna.mawi <- ggplot(dist.per.time_comp.dna.mawi, aes(x = comp, y = distance, fill = comp)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") + theme_bw() + theme(axis.title.x=element_blank(), legend.position = "none") +
  coord_cartesian(ylim = c(0, 0.27)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

```{r}

# T0 vs T3, Frozen, DNA, Jaccard

st.physeq.dna.t0_3.frozen <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '3')
dist.dna.t0_3.frozen <- phyloseq::distance(physeq = st.physeq.dna.t0_3.frozen, method = 'jaccard')
dist.dna.t0_3.frozen.df <- 
  matrix(dist.dna.t0_3.frozen)
dist.dna.t0_3.frozen.list <- reshape2::melt(dist.dna.t0_3.frozen.df)[reshape2::melt(upper.tri(dist.dna.t0_3.frozen.df))$value,]
names(dist.dna.t0_3.frozen.list) <- c("c1", "c2", "distance")
dist.dna.t0_3.frozen.list <- cbind(dist.dna.t0_3.frozen.list, "comp" = "T0_vs_T03", "type" = "Frozen")

# T0 vs T7, Frozen, DNA, Jaccard

st.physeq.dna.t0_7.frozen <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '7')
dist.dna.t0_7.frozen <- phyloseq::distance(physeq = st.physeq.dna.t0_7.frozen, method = 'jaccard')
dist.dna.t0_7.frozen.df <- 
  matrix(dist.dna.t0_7.frozen)
dist.dna.t0_7.frozen.list <- reshape2::melt(dist.dna.t0_7.frozen.df)[reshape2::melt(upper.tri(dist.dna.t0_7.frozen.df))$value,]
names(dist.dna.t0_7.frozen.list) <- c("c1", "c2", "distance")
dist.dna.t0_7.frozen.list <- cbind(dist.dna.t0_7.frozen.list, "comp" = "T0_vs_T07", "type" = "Frozen")

# T0 vs T14, Frozen, DNA, Jaccard

st.physeq.dna.t0_14.frozen <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '14')
dist.dna.t0_14.frozen <- phyloseq::distance(physeq = st.physeq.dna.t0_14.frozen, method = 'jaccard')
dist.dna.t0_14.frozen.df <- 
  matrix(dist.dna.t0_14.frozen)
dist.dna.t0_14.frozen.list <- reshape2::melt(dist.dna.t0_14.frozen.df)[reshape2::melt(upper.tri(dist.dna.t0_14.frozen.df))$value,]
names(dist.dna.t0_14.frozen.list) <- c("c1", "c2", "distance")
dist.dna.t0_14.frozen.list <- cbind(dist.dna.t0_14.frozen.list, "comp" = "T0_vs_T14", "type" = "Frozen")

# T0 vs T21, Frozen, DNA, Jaccard

st.physeq.dna.t0_21.frozen <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '21')
dist.dna.t0_21.frozen <- phyloseq::distance(physeq = st.physeq.dna.t0_21.frozen, method = 'jaccard')
dist.dna.t0_21.frozen.df <- 
  matrix(dist.dna.t0_21.frozen)
dist.dna.t0_21.frozen.list <- reshape2::melt(dist.dna.t0_21.frozen.df)[reshape2::melt(upper.tri(dist.dna.t0_21.frozen.df))$value,]
names(dist.dna.t0_21.frozen.list) <- c("c1", "c2", "distance")
dist.dna.t0_21.frozen.list <- cbind(dist.dna.t0_21.frozen.list, "comp" = "T0_vs_T21", "type" = "Frozen")

# T0 vs T40, Frozen, DNA, Jaccard

st.physeq.dna.t0_40.frozen <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '40')
dist.dna.t0_40.frozen <- phyloseq::distance(physeq = st.physeq.dna.t0_40.frozen, method = 'jaccard')
dist.dna.t0_40.frozen.df <- 
  matrix(dist.dna.t0_40.frozen)
dist.dna.t0_40.frozen.list <- reshape2::melt(dist.dna.t0_40.frozen.df)[reshape2::melt(upper.tri(dist.dna.t0_40.frozen.df))$value,]
names(dist.dna.t0_40.frozen.list) <- c("c1", "c2", "distance")
dist.dna.t0_40.frozen.list <- cbind(dist.dna.t0_40.frozen.list, "comp" = "T0_vs_T40", "type" = "Frozen")

#combine all tables
dist.per.time_comp.dna.frozen <- rbind(dist.dna.t0_3.frozen.list, dist.dna.t0_7.frozen.list, dist.dna.t0_14.frozen.list, dist.dna.t0_21.frozen.list, dist.dna.t0_40.frozen.list)

#plot
boxplot.dna.frozen <- ggplot(dist.per.time_comp.dna.frozen, aes(x = comp, y = distance, fill = comp)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") + theme_bw() + theme(axis.title.x=element_blank(), legend.position = "none") +
  coord_cartesian(ylim = c(0, 0.27)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
  
```

```{r}
dist.per.time_comp.dna <- rbind(dist.per.time_comp.dna.frozen, dist.per.time_comp.dna.mawi)

boxplot.dna.all <- ggplot(dist.per.time_comp.dna, aes(x = comp, y = distance, fill = comp, )) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") +
  facet_grid(. ~ type) +
  theme_bw() +
  theme(axis.title.x = element_blank(), legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  ggtitle("Distance Across Time Points (Jaccard)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
boxplot.dna.all
ggsave("boxplot_across_time_points (jaccard) .png",
       width = 6,
       height = 4,
       units = "in")
```

Mawi RNA

```{r}

# T0 vs T3, Mawi, RNA

st.physeq.rna.t0_3.mawi <- subset_samples(st.physeq.rna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '3')
dist.rna.t0_3.mawi <- phyloseq::distance(physeq = st.physeq.rna.t0_3.mawi, method = 'bray')
dist.rna.t0_3.mawi.df <- 
  matrix(dist.rna.t0_3.mawi)
dist.rna.t0_3.mawi.list <- reshape2::melt(dist.rna.t0_3.mawi.df)[reshape2::melt(upper.tri(dist.rna.t0_3.mawi.df))$value,]
names(dist.rna.t0_3.mawi.list) <- c("c1", "c2", "distance")
dist.rna.t0_3.mawi.list <- cbind(dist.rna.t0_3.mawi.list, "comp" = "T0_vs_T03")

# T0 vs T7, Mawi, RNA

st.physeq.rna.t0_7.mawi <- subset_samples(st.physeq.rna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '7')
dist.rna.t0_7.mawi <- phyloseq::distance(physeq = st.physeq.rna.t0_7.mawi, method = 'bray')
dist.rna.t0_7.mawi.df <- 
  matrix(dist.rna.t0_7.mawi)
dist.rna.t0_7.mawi.list <- reshape2::melt(dist.rna.t0_7.mawi.df)[reshape2::melt(upper.tri(dist.rna.t0_7.mawi.df))$value,]
names(dist.rna.t0_7.mawi.list) <- c("c1", "c2", "distance")
dist.rna.t0_7.mawi.list <- cbind(dist.rna.t0_7.mawi.list, "comp" = "T0_vs_T07")

# T0 vs T14, Mawi, RNA

st.physeq.rna.t0_14.mawi <- subset_samples(st.physeq.rna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '14')
dist.rna.t0_14.mawi <- phyloseq::distance(physeq = st.physeq.rna.t0_14.mawi, method = 'bray')
dist.rna.t0_14.mawi.df <- 
  matrix(dist.rna.t0_14.mawi)
dist.rna.t0_14.mawi.list <- reshape2::melt(dist.rna.t0_14.mawi.df)[reshape2::melt(upper.tri(dist.rna.t0_14.mawi.df))$value,]
names(dist.rna.t0_14.mawi.list) <- c("c1", "c2", "distance")
dist.rna.t0_14.mawi.list <- cbind(dist.rna.t0_14.mawi.list, "comp" = "T0_vs_T14")

# T0 vs T21, Mawi, RNA

st.physeq.rna.t0_21.mawi <- subset_samples(st.physeq.rna.r, Type == 'Mawi' & Day == '0' | Type == 'Mawi' & Day == '21')
dist.rna.t0_21.mawi <- phyloseq::distance(physeq = st.physeq.rna.t0_21.mawi, method = 'bray')
dist.rna.t0_21.mawi.df <- 
  matrix(dist.rna.t0_21.mawi)
dist.rna.t0_21.mawi.list <- reshape2::melt(dist.rna.t0_21.mawi.df)[reshape2::melt(upper.tri(dist.rna.t0_21.mawi.df))$value,]
names(dist.rna.t0_21.mawi.list) <- c("c1", "c2", "distance")
dist.rna.t0_21.mawi.list <- cbind(dist.rna.t0_21.mawi.list, "comp" = "T0_vs_T21")

#combine all tables
dist.per.time_comp.rna.mawi <- rbind(dist.rna.t0_3.mawi.list, dist.rna.t0_7.mawi.list, dist.rna.t0_14.mawi.list, dist.rna.t0_21.mawi.list)

#plot
boxplot.rna.mawi <- ggplot(dist.per.time_comp.rna.mawi, aes(x = comp, y = distance, fill = comp)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") + theme_bw() + theme(axis.title.x=element_blank(), legend.position = "none") +
  coord_cartesian(ylim = c(0, 1.02)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

```{r}

# T0 vs T3, Frozen, RNA

st.physeq.rna.t0_3.frozen <- subset_samples(st.physeq.rna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '3')
dist.rna.t0_3.frozen <- phyloseq::distance(physeq = st.physeq.rna.t0_3.frozen, method = 'bray')
dist.rna.t0_3.frozen.df <- 
  matrix(dist.rna.t0_3.frozen)
dist.rna.t0_3.frozen.list <- reshape2::melt(dist.rna.t0_3.frozen.df)[reshape2::melt(upper.tri(dist.rna.t0_3.frozen.df))$value,]
names(dist.rna.t0_3.frozen.list) <- c("c1", "c2", "distance")
dist.rna.t0_3.frozen.list <- cbind(dist.rna.t0_3.frozen.list, "comp" = "T0_vs_T03", "type" = "Frozen")

# T0 vs T7, Frozen, RNA

st.physeq.rna.t0_7.frozen <- subset_samples(st.physeq.rna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '7')
dist.rna.t0_7.frozen <- phyloseq::distance(physeq = st.physeq.rna.t0_7.frozen, method = 'bray')
dist.rna.t0_7.frozen.df <- 
  matrix(dist.rna.t0_7.mawi)
dist.rna.t0_7.frozen.list <- reshape2::melt(dist.rna.t0_7.frozen.df)[reshape2::melt(upper.tri(dist.rna.t0_7.frozen.df))$value,]
names(dist.rna.t0_7.frozen.list) <- c("c1", "c2", "distance")
dist.rna.t0_7.frozen.list <- cbind(dist.rna.t0_7.frozen.list, "comp" = "T0_vs_T07", "type" = "Frozen")

# T0 vs T14, Frozen, RNA

st.physeq.rna.t0_14.frozen <- subset_samples(st.physeq.rna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '14')
dist.rna.t0_14.frozen <- phyloseq::distance(physeq = st.physeq.rna.t0_14.frozen, method = 'bray')
dist.rna.t0_14.frozen.df <- 
  matrix(dist.rna.t0_14.frozen)
dist.rna.t0_14.frozen.list <- reshape2::melt(dist.rna.t0_14.frozen.df)[reshape2::melt(upper.tri(dist.rna.t0_14.frozen.df))$value,]
names(dist.rna.t0_14.frozen.list) <- c("c1", "c2", "distance")
dist.rna.t0_14.frozen.list <- cbind(dist.rna.t0_14.frozen.list, "comp" = "T0_vs_T14", "type" = "Frozen")

# T0 vs T21, Frozen, RNA

st.physeq.rna.t0_21.frozen <- subset_samples(st.physeq.rna.r, Type == 'Frozen' & Day == '0' | Type == 'Frozen' & Day == '21')
dist.rna.t0_21.frozen <- phyloseq::distance(physeq = st.physeq.rna.t0_21.frozen, method = 'bray')
dist.rna.t0_21.frozen.df <- 
  matrix(dist.rna.t0_21.frozen)
dist.rna.t0_21.frozen.list <- reshape2::melt(dist.rna.t0_21.frozen.df)[reshape2::melt(upper.tri(dist.rna.t0_21.frozen.df))$value,]
names(dist.rna.t0_21.frozen.list) <- c("c1", "c2", "distance")
dist.rna.t0_21.frozen.list <- cbind(dist.rna.t0_21.frozen.list, "comp" = "T0_vs_T21", "type" = "Frozen")

#combine all tables
dist.per.time_comp.rna.frozen <- rbind(dist.rna.t0_3.frozen.list, dist.rna.t0_7.frozen.list, dist.rna.t0_14.frozen.list, dist.rna.t0_21.frozen.list)

#plot
boxplot.rna.frozen <- ggplot(dist.per.time_comp.rna.frozen, aes(x = comp, y = distance, fill = comp)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") + theme_bw() + theme(axis.title.x=element_blank(), legend.position = "none") +
  coord_cartesian(ylim = c(0, 1.02)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

RNA comparison boxplots

```{r}
plot_grid(boxplot.rna.mawi, boxplot.rna.frozen, labels = c("Mawi", "Frozen"), label_size = 8)
```

Replicate Comparison (Bray-Curtis)

```{r}

# T0, Mawi, DNA, Bray

st.physeq.dna.t0.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0')
dist.dna.t0.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t0.mawi_reps, method = 'bray')
dist.dna.t0.mawi_reps.df <- 
  matrix(dist.dna.t0.mawi_reps)
dist.dna.t0.mawi_reps.list <- reshape2::melt(dist.dna.t0.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t0.mawi_reps.df))$value,]
names(dist.dna.t0.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t0.mawi_reps.list <- cbind(dist.dna.t0.mawi_reps.list, "comp" = "T0", "type" = "Mawi")

# T3, Mawi, DNA, Bray

st.physeq.dna.t3.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '3')
dist.dna.t3.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t3.mawi_reps, method = 'bray')
dist.dna.t3.mawi_reps.df <- 
  matrix(dist.dna.t3.mawi_reps)
dist.dna.t3.mawi_reps.list <- reshape2::melt(dist.dna.t3.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t3.mawi_reps.df))$value,]
names(dist.dna.t3.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t3.mawi_reps.list <- cbind(dist.dna.t3.mawi_reps.list, "comp" = "T03", "type" = "Mawi")

# T7, Mawi, DNA, Bray

st.physeq.dna.t7.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '7')
dist.dna.t7.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t7.mawi_reps, method = 'bray')
dist.dna.t7.mawi_reps.df <- 
  matrix(dist.dna.t7.mawi_reps)
dist.dna.t7.mawi_reps.list <- reshape2::melt(dist.dna.t7.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t7.mawi_reps.df))$value,]
names(dist.dna.t7.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t7.mawi_reps.list <- cbind(dist.dna.t7.mawi_reps.list, "comp" = "T07", "type" = "Mawi")

# T14, Mawi, DNA, Bray

st.physeq.dna.t14.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '14')
dist.dna.t14.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t14.mawi_reps, method = 'bray')
dist.dna.t14.mawi_reps.df <- 
  matrix(dist.dna.t14.mawi_reps)
dist.dna.t14.mawi_reps.list <- reshape2::melt(dist.dna.t14.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t14.mawi_reps.df))$value,]
names(dist.dna.t14.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t14.mawi_reps.list <- cbind(dist.dna.t14.mawi_reps.list, "comp" = "T14", "type" = "Mawi")

# T21, Mawi, DNA, Bray

st.physeq.dna.t21.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '21')
dist.dna.t21.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t21.mawi_reps, method = 'bray')
dist.dna.t21.mawi_reps.df <- 
  matrix(dist.dna.t21.mawi_reps)
dist.dna.t21.mawi_reps.list <- reshape2::melt(dist.dna.t21.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t21.mawi_reps.df))$value,]
names(dist.dna.t21.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t21.mawi_reps.list <- cbind(dist.dna.t21.mawi_reps.list, "comp" = "T21", "type" = "Mawi")

# T40, Mawi, DNA, Bray

st.physeq.dna.t40.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '40')
dist.dna.t40.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t40.mawi_reps, method = 'bray')
dist.dna.t40.mawi_reps.df <- 
  matrix(dist.dna.t40.mawi_reps)
dist.dna.t40.mawi_reps.list <- reshape2::melt(dist.dna.t40.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t40.mawi_reps.df))$value,]
names(dist.dna.t40.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t40.mawi_reps.list <- cbind(dist.dna.t40.mawi_reps.list, "comp" = "T40", "type" = "Mawi")

#combine all tables
dist.per.time_comp.dna.mawi_reps <- rbind(dist.dna.t0.mawi_reps.list, dist.dna.t3.mawi_reps.list, dist.dna.t7.mawi_reps.list, dist.dna.t14.mawi_reps.list, dist.dna.t21.mawi_reps.list, dist.dna.t40.mawi_reps.list)

#plot
boxplot.mawi_reps.dna <- ggplot(dist.per.time_comp.dna.mawi_reps, aes(x = comp, y = distance, fill = comp)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") + theme_bw() + theme(axis.title.x=element_blank(), legend.position = "none") +
  coord_cartesian(ylim = c(0, 0.25))
```

```{r}

# T0, Frozen, DNA, Bray

st.physeq.dna.t0.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0')
dist.dna.t0.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t0.frozen_reps, method = 'bray')
dist.dna.t0.frozen_reps.df <- 
  matrix(dist.dna.t0.frozen_reps)
dist.dna.t0.frozen_reps.list <- reshape2::melt(dist.dna.t0.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t0.frozen_reps.df))$value,]
names(dist.dna.t0.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t0.frozen_reps.list <- cbind(dist.dna.t0.frozen_reps.list, "comp" = "T0", "type" = "Frozen")

# T3, Frozen, DNA, Bray

st.physeq.dna.t3.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '3')
dist.dna.t3.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t3.frozen_reps, method = 'bray')
dist.dna.t3.frozen_reps.df <- 
  matrix(dist.dna.t3.frozen_reps)
dist.dna.t3.frozen_reps.list <- reshape2::melt(dist.dna.t3.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t3.frozen_reps.df))$value,]
names(dist.dna.t3.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t3.frozen_reps.list <- cbind(dist.dna.t3.frozen_reps.list, "comp" = "T03", "type" = "Frozen")

# T7, Frozen, DNA, Bray

st.physeq.dna.t7.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '7')
dist.dna.t7.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t7.frozen_reps, method = 'bray')
dist.dna.t7.frozen_reps.df <- 
  matrix(dist.dna.t7.frozen_reps)
dist.dna.t7.frozen_reps.list <- reshape2::melt(dist.dna.t7.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t7.frozen_reps.df))$value,]
names(dist.dna.t7.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t7.frozen_reps.list <- cbind(dist.dna.t7.frozen_reps.list, "comp" = "T07", "type" = "Frozen")

# T14, Frozen, DNA, Bray

st.physeq.dna.t14.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '14')
dist.dna.t14.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t14.frozen_reps, method = 'bray')
dist.dna.t14.frozen_reps.df <- 
  matrix(dist.dna.t14.frozen_reps)
dist.dna.t14.frozen_reps.list <- reshape2::melt(dist.dna.t14.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t14.frozen_reps.df))$value,]
names(dist.dna.t14.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t14.frozen_reps.list <- cbind(dist.dna.t14.frozen_reps.list, "comp" = "T14", "type" = "Frozen")

# T21, Frozen, DNA, Bray

st.physeq.dna.t21.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '21')
dist.dna.t21.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t21.frozen_reps, method = 'bray')
dist.dna.t21.frozen_reps.df <- 
  matrix(dist.dna.t21.frozen_reps)
dist.dna.t21.frozen_reps.list <- reshape2::melt(dist.dna.t21.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t21.frozen_reps.df))$value,]
names(dist.dna.t21.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t21.frozen_reps.list <- cbind(dist.dna.t21.frozen_reps.list, "comp" = "T21", "type" = "Frozen")

# T40, Frozen, DNA, Bray

st.physeq.dna.t40.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '40')
dist.dna.t40.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t40.frozen_reps, method = 'bray')
dist.dna.t40.frozen_reps.df <- 
  matrix(dist.dna.t40.frozen_reps)
dist.dna.t40.frozen_reps.list <- reshape2::melt(dist.dna.t40.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t40.frozen_reps.df))$value,]
names(dist.dna.t40.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t40.frozen_reps.list <- cbind(dist.dna.t40.frozen_reps.list, "comp" = "T40", "type" = "Frozen")

#combine all tables
dist.per.time_comp.dna.frozen_reps <- rbind(dist.dna.t0.frozen_reps.list, dist.dna.t3.frozen_reps.list, dist.dna.t7.frozen_reps.list, dist.dna.t14.frozen_reps.list, dist.dna.t21.frozen_reps.list, dist.dna.t40.frozen_reps.list)

#plot
boxplot.frozen_reps.dna <- ggplot(dist.per.time_comp.dna.frozen_reps, aes(x = comp, y = distance, fill = comp)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") + theme_bw() + theme(axis.title.x=element_blank(), legend.position = "none") +
  coord_cartesian(ylim = c(0, 0.25))
```

```{r}
dist.per.reps.dna <- rbind(dist.per.time_comp.dna.frozen_reps, dist.per.time_comp.dna.mawi_reps)

boxplot.reps.dna <- ggplot(dist.per.reps.dna, aes(x = comp, y = distance, fill = comp, )) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") +
  facet_grid(. ~ type) +
  theme_bw() +
  theme(axis.title.x = element_blank(), legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  ggtitle("Distance Comparison By Replicate (Bray-Curtis)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
boxplot.reps.dna
ggsave("boxplot_reps (bray) .png",
       width = 6,
       height = 4,
       units = "in")
```
Replicate Comparison (Jaccard)

```{r}

# T0, Mawi, DNA, Jaccard

st.physeq.dna.t0.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '0')
dist.dna.t0.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t0.mawi_reps, method = 'jaccard')
dist.dna.t0.mawi_reps.df <- 
  matrix(dist.dna.t0.mawi_reps)
dist.dna.t0.mawi_reps.list <- reshape2::melt(dist.dna.t0.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t0.mawi_reps.df))$value,]
names(dist.dna.t0.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t0.mawi_reps.list <- cbind(dist.dna.t0.mawi_reps.list, "comp" = "T0", "type" = "Mawi")

# T3, Mawi, DNA, Jaccard

st.physeq.dna.t3.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '3')
dist.dna.t3.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t3.mawi_reps, method = 'jaccard')
dist.dna.t3.mawi_reps.df <- 
  matrix(dist.dna.t3.mawi_reps)
dist.dna.t3.mawi_reps.list <- reshape2::melt(dist.dna.t3.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t3.mawi_reps.df))$value,]
names(dist.dna.t3.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t3.mawi_reps.list <- cbind(dist.dna.t3.mawi_reps.list, "comp" = "T03", "type" = "Mawi")

# T7, Mawi, DNA, Jaccard

st.physeq.dna.t7.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '7')
dist.dna.t7.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t7.mawi_reps, method = 'jaccard')
dist.dna.t7.mawi_reps.df <- 
  matrix(dist.dna.t7.mawi_reps)
dist.dna.t7.mawi_reps.list <- reshape2::melt(dist.dna.t7.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t7.mawi_reps.df))$value,]
names(dist.dna.t7.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t7.mawi_reps.list <- cbind(dist.dna.t7.mawi_reps.list, "comp" = "T07", "type" = "Mawi")

# T14, Mawi, DNA, Jaccard

st.physeq.dna.t14.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '14')
dist.dna.t14.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t14.mawi_reps, method = 'jaccard')
dist.dna.t14.mawi_reps.df <- 
  matrix(dist.dna.t14.mawi_reps)
dist.dna.t14.mawi_reps.list <- reshape2::melt(dist.dna.t14.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t14.mawi_reps.df))$value,]
names(dist.dna.t14.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t14.mawi_reps.list <- cbind(dist.dna.t14.mawi_reps.list, "comp" = "T14", "type" = "Mawi")

# T21, Mawi, DNA, Jaccard

st.physeq.dna.t21.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '21')
dist.dna.t21.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t21.mawi_reps, method = 'jaccard')
dist.dna.t21.mawi_reps.df <- 
  matrix(dist.dna.t21.mawi_reps)
dist.dna.t21.mawi_reps.list <- reshape2::melt(dist.dna.t21.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t21.mawi_reps.df))$value,]
names(dist.dna.t21.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t21.mawi_reps.list <- cbind(dist.dna.t21.mawi_reps.list, "comp" = "T21", "type" = "Mawi")

# T40, Mawi, DNA, Jaccard

st.physeq.dna.t40.mawi_reps <- subset_samples(st.physeq.dna.r, Type == 'Mawi' & Day == '40')
dist.dna.t40.mawi_reps <- phyloseq::distance(physeq = st.physeq.dna.t40.mawi_reps, method = 'jaccard')
dist.dna.t40.mawi_reps.df <- 
  matrix(dist.dna.t40.mawi_reps)
dist.dna.t40.mawi_reps.list <- reshape2::melt(dist.dna.t40.mawi_reps.df)[reshape2::melt(upper.tri(dist.dna.t40.mawi_reps.df))$value,]
names(dist.dna.t40.mawi_reps.list) <- c("c1", "c2", "distance")
dist.dna.t40.mawi_reps.list <- cbind(dist.dna.t40.mawi_reps.list, "comp" = "T40", "type" = "Mawi")

#combine all tables
dist.per.time_comp.dna.mawi_reps <- rbind(dist.dna.t0.mawi_reps.list, dist.dna.t3.mawi_reps.list, dist.dna.t7.mawi_reps.list, dist.dna.t14.mawi_reps.list, dist.dna.t21.mawi_reps.list, dist.dna.t40.mawi_reps.list)

#plot
boxplot.mawi_reps.dna <- ggplot(dist.per.time_comp.dna.mawi_reps, aes(x = comp, y = distance, fill = comp)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") + theme_bw() + theme(axis.title.x=element_blank(), legend.position = "none") +
  coord_cartesian(ylim = c(0, 0.25))
```

```{r}

# T0, Frozen, DNA, Jaccard

st.physeq.dna.t0.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '0')
dist.dna.t0.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t0.frozen_reps, method = 'jaccard')
dist.dna.t0.frozen_reps.df <- 
  matrix(dist.dna.t0.frozen_reps)
dist.dna.t0.frozen_reps.list <- reshape2::melt(dist.dna.t0.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t0.frozen_reps.df))$value,]
names(dist.dna.t0.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t0.frozen_reps.list <- cbind(dist.dna.t0.frozen_reps.list, "comp" = "T0", "type" = "Frozen")

# T3, Frozen, DNA, Jaccard

st.physeq.dna.t3.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '3')
dist.dna.t3.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t3.frozen_reps, method = 'jaccard')
dist.dna.t3.frozen_reps.df <- 
  matrix(dist.dna.t3.frozen_reps)
dist.dna.t3.frozen_reps.list <- reshape2::melt(dist.dna.t3.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t3.frozen_reps.df))$value,]
names(dist.dna.t3.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t3.frozen_reps.list <- cbind(dist.dna.t3.frozen_reps.list, "comp" = "T03", "type" = "Frozen")

# T7, Frozen, DNA, Jaccard

st.physeq.dna.t7.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '7')
dist.dna.t7.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t7.frozen_reps, method = 'jaccard')
dist.dna.t7.frozen_reps.df <- 
  matrix(dist.dna.t7.frozen_reps)
dist.dna.t7.frozen_reps.list <- reshape2::melt(dist.dna.t7.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t7.frozen_reps.df))$value,]
names(dist.dna.t7.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t7.frozen_reps.list <- cbind(dist.dna.t7.frozen_reps.list, "comp" = "T07", "type" = "Frozen")

# T14, Frozen, DNA, Jaccard

st.physeq.dna.t14.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '14')
dist.dna.t14.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t14.frozen_reps, method = 'jaccard')
dist.dna.t14.frozen_reps.df <- 
  matrix(dist.dna.t14.frozen_reps)
dist.dna.t14.frozen_reps.list <- reshape2::melt(dist.dna.t14.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t14.frozen_reps.df))$value,]
names(dist.dna.t14.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t14.frozen_reps.list <- cbind(dist.dna.t14.frozen_reps.list, "comp" = "T14", "type" = "Frozen")

# T21, Frozen, DNA, Jaccard

st.physeq.dna.t21.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '21')
dist.dna.t21.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t21.frozen_reps, method = 'jaccard')
dist.dna.t21.frozen_reps.df <- 
  matrix(dist.dna.t21.frozen_reps)
dist.dna.t21.frozen_reps.list <- reshape2::melt(dist.dna.t21.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t21.frozen_reps.df))$value,]
names(dist.dna.t21.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t21.frozen_reps.list <- cbind(dist.dna.t21.frozen_reps.list, "comp" = "T21", "type" = "Frozen")

# T40, Frozen, DNA, Jaccard

st.physeq.dna.t40.frozen_reps <- subset_samples(st.physeq.dna.r, Type == 'Frozen' & Day == '40')
dist.dna.t40.frozen_reps <- phyloseq::distance(physeq = st.physeq.dna.t40.frozen_reps, method = 'jaccard')
dist.dna.t40.frozen_reps.df <- 
  matrix(dist.dna.t40.frozen_reps)
dist.dna.t40.frozen_reps.list <- reshape2::melt(dist.dna.t40.frozen_reps.df)[reshape2::melt(upper.tri(dist.dna.t40.frozen_reps.df))$value,]
names(dist.dna.t40.frozen_reps.list) <- c("c1", "c2", "distance")
dist.dna.t40.frozen_reps.list <- cbind(dist.dna.t40.frozen_reps.list, "comp" = "T40", "type" = "Frozen")

#combine all tables
dist.per.time_comp.dna.frozen_reps <- rbind(dist.dna.t0.frozen_reps.list, dist.dna.t3.frozen_reps.list, dist.dna.t7.frozen_reps.list, dist.dna.t14.frozen_reps.list, dist.dna.t21.frozen_reps.list, dist.dna.t40.frozen_reps.list)

#plot
boxplot.frozen_reps.dna <- ggplot(dist.per.time_comp.dna.frozen_reps, aes(x = comp, y = distance, fill = comp)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") + theme_bw() + theme(axis.title.x=element_blank(), legend.position = "none") +
  coord_cartesian(ylim = c(0, 0.25))
```

```{r}
dist.per.reps.dna <- rbind(dist.per.time_comp.dna.frozen_reps, dist.per.time_comp.dna.mawi_reps)

boxplot.reps.dna <- ggplot(dist.per.reps.dna, aes(x = comp, y = distance, fill = comp, )) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 8,
               outlier.size = 2) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") +
  facet_grid(. ~ type) +
  theme_bw() +
  theme(axis.title.x = element_blank(), legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  ggtitle("Distance Comparison By Replicate (Jaccard)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
boxplot.reps.dna
ggsave("boxplot_reps (jaccard) .png",
       width = 6,
       height = 4,
       units = "in")
```

```{r}

st.physeq.dna.r.mawi <- subset_samples(st.physeq.dna.r, Type == "Mawi")
dna.mawi.merged <- merge_samples(st.physeq.dna.r.mawi, "Day")
dna.mawi.merged.pool <- merge_samples(st.physeq.dna.r.mawi, "Type")

#sample_data(dna.mawi.merged)$Day <- levels(sample_data(st.physeq.dna.r.mawi)$Day)

dna.mawi.merged <- transform_sample_counts(dna.mawi.merged, function(x) 100 * x/sum(x))
dna.mawi.merged.taxphylum <- tax_glom(dna.mawi.merged, "Phylum")
phylum10 <- names(sort(taxa_sums(dna.mawi.merged.taxphylum), TRUE)[1:10])
dna.mawi.phylum10 <- prune_taxa(phylum10, dna.mawi.merged.taxphylum)

title <- "Relative Abundance of Bacterial Communities in Mawi Stool Samples"
sample_data(dna.mawi.phylum10)$Day <- factor(sample_data(dna.mawi.phylum10)$Day, levels = c("0", "3", "7", "14", "21", "40"))

bp.mawi <- plot_bar(dna.mawi.phylum10, "Day", fill = "Phylum", title=title) +
  scale_fill_manual(breaks = c("Actinobacteria", "Aquificae", "Bacteria_u_p", "Bacteroidetes", "Chloroflexi", "Euryarchaeota", "Firmicutes", "Gemmatimonadetes", "Proteobacteria", "Verrucomicrobia"),
                    values = c("coral2", "cyan", "mediumpurple", "goldenrod", "darkseagreen", "darkorange", "royalblue2", "plum4", "tan3", "olivedrab3"))
                    
```

```{r}
test_subset <- subset_samples(st.physeq.dna, Type == "Mawi")
test_dna <- merge_samples(test_subset, "Day")

test_phylum <- tax_glom(test_dna, "Phylum")
phylum20 <- names(sort(taxa_sums(test_phylum), TRUE)[1:20])
test_phylum20 <- prune_taxa(phylum20, test_phylum)
```

```{r}

st.physeq.dna.r.frozen <- subset_samples(st.physeq.dna.r, Type == "Frozen")
dna.frozen.merged <- merge_samples(st.physeq.dna.r.frozen, "Day")

#sample_data(dna.mawi.merged)$Day <- levels(sample_data(st.physeq.dna.r.mawi)$Day)

dna.frozen.merged <- transform_sample_counts(dna.frozen.merged, function(x) 100 * x/sum(x))
dna.frozen.merged.taxphylum <- tax_glom(dna.frozen.merged, "Phylum")
phylum10 <- names(sort(taxa_sums(dna.frozen.merged.taxphylum), TRUE)[1:10])
dna.frozen.phylum10 <- prune_taxa(phylum10, dna.frozen.merged.taxphylum)

title <- "Relative Abundance of Bacterial Communities in Frozen Stool Samples"
sample_data(dna.frozen.phylum10)$Day <- factor(sample_data(dna.frozen.phylum10)$Day, levels = c("0", "3", "7", "14", "21", "40"))

bp.frozen <- plot_bar(dna.frozen.phylum10, "Day", fill = "Phylum", title=title) +
  scale_fill_manual(breaks = c("Acidobacteria", "Actinobacteria", "Aquificae", "Bacteria_u_p", "Bacteroidetes", "Dictyoglomi", "Euryarchaeota", "Firmicutes", "Proteobacteria", "Verrucomicrobia"),
                    values = c("lightsteelblue2", "coral2", "cyan", "mediumpurple", "goldenrod", "peachpuff1", "darkorange", "royalblue2", "tan3", "olivedrab3"))

```

```{r}
test_subset <- subset_samples(st.physeq.dna, Type == "Frozen")
test_dna <- merge_samples(test_subset, "Day")

test_phylum <- tax_glom(test_dna, "Phylum")
phylum20 <- names(sort(taxa_sums(test_phylum), TRUE)[1:20])
test_phylum20 <- prune_taxa(phylum20, test_phylum)
```
Acidobacteria = "lightsteelblue2"
Actinobacteria = "coral2"
Aquificae = "cyan"
Bacteria_u_p = "mediumpurple"
Bacteroidetes = "goldenrod"
Chloroflexi = "darkseagreen1"
Dictyoglomi = "peachpuff1"
Euryarchaeota = "darkorange"
Firmicutes = "royalblue2"
Gemmatimonadetes = "plum4"
Proteobacteria = "tan3"
Verrumicrobia = "olivedrab3"

```{r}
plot_grid(bp.mawi, bp.frozen, labels = c("Mawi", "Frozen"), label_size = 8)
```

```{r}
obj <- st.physeq.dna.r
otutable <- data.frame(OTU = rownames(phyloseq::otu_table(obj)@.Data),
                       phyloseq::otu_table(obj)@.Data,
                       phyloseq::tax_table(obj)@.Data,
                       check.names = FALSE)

metadata <- data.frame(phyloseq::sample_data(obj),
                       check.names = FALSE)

my_ampvis2_object <- amp_load(otutable, metadata)

mawi_ampvis <- amp_subset_samples(my_ampvis2_object,
                                  Type %in% c("Mawi"),
                                  normalise = FALSE)

ampvis_dna.genus <- amp_heatmap(my_ampvis2_object,
                          #group_by = "Day",
                          facet_by = "Type",
                          tax_aggregate = "Genus",
                          tax_add = c("Phylum"),
                          tax_show = 30,
                          color_vector = c("white", "darkblue"),
                          plot_colorscale = "sqrt",
                          plot_values = TRUE,
                          plot_values_size = 3) +
  theme(axis.text.x = element_text(angle = 45, size = 10, vjust = 1),
        axis.text.y = element_text(size = 10),
        legend.position = "right") +
  ggtitle("Genus Abundance By Sample") +
  theme(plot.title = element_text(hjust = 0.5))

ampvis_dna.species <- amp_heatmap(my_ampvis2_object,
                          #group_by = "Day",
                          facet_by = "Type",
                          tax_aggregate = "Species",
                          tax_add = c("Phylum"),
                          tax_show = 30,
                          color_vector = c("white", "darkblue"),
                          plot_colorscale = "sqrt",
                          plot_values = TRUE,
                          plot_values_size = 3) +
  theme(axis.text.x = element_text(angle = 45, size = 10, vjust = 1),
        axis.text.y = element_text(size = 10),
        legend.position = "right") +
  ggtitle("Species Abundance By Sample") +
  theme(plot.title = element_text(hjust = 0.5))


ampvis_day <- amp_heatmap(my_ampvis2_object,
                          group_by = "Type",
                          facet_by = "Day",
                          tax_aggregate = "Species",
                          tax_add = c("Phylum"),
                          tax_show = 30,
                          color_vector = c("white", "darkblue"),
                          plot_colorscale = "sqrt",
                          plot_values = TRUE,
                          plot_values_size = 3) +
  theme(axis.text.x = element_text(angle = 45, size = 10, vjust = 1),
        axis.text.y = element_text(size = 10),
        legend.position = "right") +
  ggtitle("Species Abundance By Sample Per Day") +
  theme(plot.title = element_text(hjust = 0.5))


ampvis_day_reps <- amp_heatmap(my_ampvis2_object,
                          #group_by = "Type",
                          facet_by = "Day",
                          tax_aggregate = "Species",
                          tax_add = c("Phylum"),
                          tax_show = 30,
                          color_vector = c("white", "darkblue"),
                          plot_colorscale = "sqrt",
                          plot_values = TRUE,
                          plot_values_size = 3) +
  theme(axis.text.x = element_text(angle = 45, size = 10, vjust = 1),
        axis.text.y = element_text(size = 10),
        legend.position = "right") +
  ggtitle("Species Abundance By Replicate Per Day") +
  theme(plot.title = element_text(hjust = 0.5))


```

```{r}
ampvis_dna.species
ggsave("ampvis_heatmap_by_species .png",
       width = 16,
       height = 9,
       units = "in")
```

```{r}
ampvis_dna.genus
ggsave("ampvis_heatmap_by_genus .png",
       width = 16,
       height = 9,
       units = "in")
```

```{r}
ampvis_day
ggsave("ampvis_heatmap_by_day .png",
       width = 16,
       height = 9,
       units = "in")
```
```{r}
ampvis_day_reps
ggsave("ampvis_heatmap_by_day_and_reps.png",
       width = 16,
       height = 9,
       units = "in")
```

```{r}
test_phylum <- tax_glom(st.physeq.dna, "Species")
phylum20 <- names(sort(taxa_sums(test_phylum), TRUE)[1:20])
test_phylum20 <- prune_taxa(phylum20, test_phylum)
```

```{r}
palette_info <- brewer.pal.info[brewer.pal.info$category == "qual", ]
palette_all <- unlist(mapply(brewer.pal,
                             palette_info$maxcolors,
                             rownames(palette_info)))

set.seed(13)
palette_bar <- sample(palette_all, 20)
```

```{r}
bar_species <- plot_bar(test_phylum20, fill = "Species") +
  theme(legend.key.size = unit(0.3, "cm")) +
  scale_fill_manual(values = palette_bar)
bar_species
```

```{r}
library(MicrobiotaProcess)
```

```{r}
phylumtaxa <- get_taxadf(obj=st.physeq.dna, taxlevel=2)
pphylum <- ggbartax(obj=phylumtaxa, facetNames="Type") +
  xlab(NULL) +
  ylab("relative abundance (%)") +
  scale_fill_manual(values = c(colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(31))) +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5))

pphylum
```

```{r}
genustaxa <- get_taxadf(obj=st.physeq.dna, taxlevel=6)
pgenus <- ggbartax(obj=genustaxa, facetNames="Type", topn = 30) +
  xlab(NULL) +
  ylab("relative abundance (%)") +
  scale_fill_manual(values = c(colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(31))) +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5)) +
  ggtitle("Genus Abundance By Sample") +
  theme(plot.title = element_text(hjust = 0.5))


pgenus
ggsave("ampvis_barchart_genus_top30.png",
       width = 16,
       height = 9,
       units = "in")
```

```{r}
speciestaxa <- get_taxadf(obj=st.physeq.dna, taxlevel=7)
pspecies <- ggbartax(obj=speciestaxa, facetNames="Type", topn = 30) +
  xlab(NULL) +
  ylab("relative abundance (%)") +
  scale_fill_manual(values = c(colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(31))) +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5)) +
  ggtitle("Species Abundance By Sample") +
  theme(plot.title = element_text(hjust = 0.5))


pspecies
ggsave("ampvis_barchart_species_top30.png",
       width = 16,
       height = 9,
       units = "in")
```



```{r}
library(ggtree)
hcsample <- get_clust(obj = st.physeq.dna, distmethod = "bray",
                      method = "hellinger",
                      hclustmethod = "average")

cplot <- ggclust(obj = hcsample,
                 layout ="rectangular",
                 pointsize = 1,
                 fontsize = 1,
                 factorNames = c("Type")
                 ) +
  scale_color_manual(values = c("#00AED7", "#FD9347")) +
  theme_tree2(legend.position = "right",
              plot.title = element_text(face = "bold", lineheight = 25, hjust = 0.5))
```

```{r}
cplot
ggsave("cluster_all_samples_dna.png",
       width = 16,
       height = 9,
       units = "in")
```

```{r}
detach("package:MicrobiotaProcess", unload = TRUE)
```

```{r}
TypeCategories = levels(get_variable(st.physeq.dna.r, "Day"))
sample_data(st.physeq.dna.r)$Day <- factor(sample_data(st.physeq.dna.r)$Day, levels = c("0", "3", "7", "14", "21", "40"))

richness.dna.invs <- plot_richness(st.physeq.dna.r,
                              measures = "invsimpson",
                              x = "Day",
                              color = "Type",
                              shape = NA
                              ) +
  geom_boxplot() +
  geom_point(size = 0) +
  coord_cartesian(ylim = c(0, 50)) +
  scale_color_manual(values = c('Frozen' = 'blue', 'Mawi' = '#ff6600'))

richness.dna.chao <- plot_richness(st.physeq.dna.r,
                              measures = "Chao1",
                              x = "Day",
                              color = "Type",
                              shape = NA,
                              ) +
  geom_boxplot() +
  geom_point(size = 0) +
  coord_cartesian(ylim = c(0, 350)) +
  scale_color_manual(values = c('Frozen' = 'blue', 'Mawi' = '#ff6600'))

richness.dna.shannon <- plot_richness(st.physeq.dna.r,
                              measures = "Shannon",
                              x = "Day",
                              color = "Type",
                              shape = NA
                              ) +
  geom_boxplot() +
  geom_point(size = 0) +
  #coord_cartesian(ylim = c(0, 4.2)) +
  scale_color_manual(values = c('Frozen' = 'blue', 'Mawi' = '#ff6600'))
```

```{r}
richness.dna.invs
ggsave("richness_all_samples_dna_invsimpson.png",
       width = 6,
       height = 4,
       units = "in")

richness.dna.chao
ggsave("richness_all_samples_dna_chao.png",
       width = 6,
       height = 4,
       units = "in")

richness.dna.shannon
ggsave("richness_all_samples_dna_shannon_2.png",
       width = 6,
       height = 4,
       units = "in")
```
Venn Diagram of initial OTU filter
```{r}
vennp.day <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("#00AED7", "#FD9347"),
                      cat.col=c("#00AED7", "#FD9347"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

grid.draw(vennp.day)
```

```{r}
st.physeq.dna.mawi.d0 <- subset_samples(st.physeq.dna, Day == "0" & Type == "Mawi")
st.physeq.dna.mawi.d3 <- subset_samples(st.physeq.dna, Day == "3" & Type == "Mawi")
st.physeq.dna.mawi.d7 <- subset_samples(st.physeq.dna, Day == "7" & Type == "Mawi")
st.physeq.dna.mawi.d14 <- subset_samples(st.physeq.dna, Day == "14" & Type == "Mawi")
st.physeq.dna.mawi.d21 <- subset_samples(st.physeq.dna, Day == "21" & Type == "Mawi")
st.physeq.dna.mawi.d40 <- subset_samples(st.physeq.dna, Day == "40"& Type == "Mawi")

st.physeq.dna.frozen.d0 <- subset_samples(st.physeq.dna, Day == "0" & Type == "Frozen")
st.physeq.dna.frozen.d3 <- subset_samples(st.physeq.dna, Day == "3" & Type == "Frozen")
st.physeq.dna.frozen.d7 <- subset_samples(st.physeq.dna, Day == "7" & Type == "Frozen")
st.physeq.dna.frozen.d14 <- subset_samples(st.physeq.dna, Day == "14" & Type == "Frozen")
st.physeq.dna.frozen.d21 <- subset_samples(st.physeq.dna, Day == "21" & Type == "Frozen")
st.physeq.dna.frozen.d40 <- subset_samples(st.physeq.dna, Day == "40"& Type == "Frozen")

library(MicrobiotaProcess)
vennlist <- get_vennlist(obj=st.physeq.dna.mawi.d0, factorNames="Pool")
ItemsList.mawi.d0 <- venn(vennlist, show.plot = FALSE)
intersections.mawi.d0 <- lengths(attributes(ItemsList.mawi.d0)$intersections)
matrix.mawi.d0 <- t(
  matrix(intersections.mawi.d0))
vennp.mawi.d0 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.mawi.d3, factorNames="Pool")
ItemsList.mawi.d3 <- venn(vennlist, show.plot = FALSE)
intersections.mawi.d3 <- lengths(attributes(ItemsList.mawi.d3)$intersections)
matrix.mawi.d3 <- t(
  matrix(intersections.mawi.d3))
vennp.mawi.d3 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.mawi.d7, factorNames="Pool")
ItemsList.mawi.d7 <- venn(vennlist, show.plot = FALSE)
intersections.mawi.d7 <- lengths(attributes(ItemsList.mawi.d7)$intersections)
matrix.mawi.d7 <- t(
  matrix(intersections.mawi.d7))
vennp.mawi.d7 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.mawi.d14, factorNames="Pool")
ItemsList.mawi.d14 <- venn(vennlist, show.plot = FALSE)
intersections.mawi.d14 <- lengths(attributes(ItemsList.mawi.d14)$intersections)
matrix.mawi.d14 <- t(
  matrix(intersections.mawi.d14))
vennp.mawi.d14 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.mawi.d21, factorNames="Pool")
ItemsList.mawi.d21 <- venn(vennlist, show.plot = FALSE)
intersections.mawi.d21 <- lengths(attributes(ItemsList.mawi.d21)$intersections)
matrix.mawi.d21 <- t(
  matrix(intersections.mawi.d21))
vennp.mawi.d21 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.mawi.d40, factorNames="Pool")
ItemsList.mawi.d40 <- venn(vennlist, show.plot = FALSE)
intersections.mawi.d40 <- lengths(attributes(ItemsList.mawi.d40)$intersections)
matrix.mawi.d40 <- t(
  matrix(intersections.mawi.d40))
vennp.mawi.d40 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.frozen.d0, factorNames="Pool")
ItemsList.frozen.d0 <- venn(vennlist, show.plot = FALSE)
intersections.frozen.d0 <- lengths(attributes(ItemsList.frozen.d0)$intersections)
matrix.frozen.d0 <- t(
  matrix(intersections.frozen.d0))
vennp.frozen.d0 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.frozen.d3, factorNames="Pool")
ItemsList.frozen.d3 <- venn(vennlist, show.plot = FALSE)
intersections.frozen.d3 <- lengths(attributes(ItemsList.frozen.d3)$intersections)
matrix.frozen.d3 <- t(
  matrix(intersections.frozen.d3))
vennp.frozen.d3 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.frozen.d7, factorNames="Pool")
ItemsList.frozen.d7 <- venn(vennlist, show.plot = FALSE)
intersections.frozen.d7 <- lengths(attributes(ItemsList.frozen.d7)$intersections)
matrix.frozen.d7 <- t(
  matrix(intersections.frozen.d7))
vennp.frozen.d7 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.frozen.d14, factorNames="Pool")
ItemsList.frozen.d14 <- venn(vennlist, show.plot = FALSE)
intersections.frozen.d14 <- lengths(attributes(ItemsList.frozen.d14)$intersections)
matrix.frozen.d14 <- t(
  matrix(intersections.frozen.d14))
vennp.frozen.d14 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.frozen.d21, factorNames="Pool")
ItemsList.frozen.d21 <- venn(vennlist, show.plot = FALSE)
intersections.frozen.d21 <- lengths(attributes(ItemsList.frozen.d21)$intersections)
matrix.frozen.d21 <- t(
  matrix(intersections.frozen.d21))
vennp.frozen.d21 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")

vennlist <- get_vennlist(obj=st.physeq.dna.frozen.d40, factorNames="Pool")
ItemsList.frozen.d40 <- venn(vennlist, show.plot = FALSE)
intersections.frozen.d40 <- lengths(attributes(ItemsList.frozen.d40)$intersections)
matrix.frozen.d40 <- t(
  matrix(intersections.frozen.d40))
vennp.frozen.d40 <- venn.diagram(vennlist,
                      height=5,
                      width=5, 
                      filename=NULL, 
                      fill=c("orange", "blue", "green"),
                      cat.col=c("orange", "blue", "green"),
                      alpha = 0.85, 
                      fontfamily = "serif",
                      fontface = "bold",
                      cex = 1.2,
                      cat.cex = 1.3,
                      cat.default.pos = "outer",
                      cat.dist=0.1,
                      margin = 0.1, 
                      lwd = 3,
                      lty ='dotted',
                      imagetype = "svg")
```

```{r}
#extract list of elements from the venn diagram
all_intersections <- rbind(matrix.mawi.d0, matrix.mawi.d3, matrix.mawi.d7, matrix.mawi.d14, matrix.mawi.d21, matrix.mawi.d40, matrix.frozen.d0, matrix.frozen.d3, matrix.frozen.d7, matrix.frozen.d14, matrix.frozen.d21, matrix.frozen.d40)
all_intersections <- 
  data.frame(all_intersections)
rownames(all_intersections) <- c("Mawi_D0", "Mawi_D03", "Mawi_D07", "Mawi_D14", "Mawi_D21", "Mawi_D40", "Frozen_D0", "Frozen_D03", "Frozen_D07", "Frozen_D14", "Frozen_D21", "Frozen_D40")
all_intersections

all_intersections <- tibble::rownames_to_column(all_intersections, "Sample_Day")
ggplot(data = all_intersections, aes(x = "Sample_Day", y = all_intersections[, 2:7]))
```

####### Use this ####### 
#So, remove samples with less than 10000 reads, depth = 10000
#D3.S2 (2), D1.S2 (417) were excluded
#Next size is 9923 for D3.S3
Succ_cl.physeqFf = prune_samples(sample_sums(Succ_cl.physeqF2)>=9900, Succ_cl.physeqF2)
Succ_cl.physeqFf #4229 taxa and 84 samples #12661 (w/o sample instances)

#### Rarefied and filtered OTU table ####
##############################################
# Initialize matrices to store richness and evenness estimates
```{r}
nsamp = nsamples(st.physeq.dna)
trials = 100
```

```{r}
richness <- matrix(nrow = nsamp, ncol = trials)
row.names(richness) <- sample_names(st.physeq.dna)

evenness <- matrix(nrow = nsamp, ncol = trials)
row.names(evenness) <- sample_names(st.physeq.dna)
```

```{r}
# Set seed for reproducibility
set.seed(13)
for (i in 1:100) {
  # Subsample 100 times and trim OTUs with total nseq=0
  st.physeq.dna.rar <- rarefy_even_depth(st.physeq.dna, sample.size = 30000, verbose = FALSE, replace = TRUE, trimOTUs = TRUE)
  
  # Calculate richness
  rich <- 
    numeric(
    matrix(estimate_richness(st.physeq.dna.rar, measures = "Observed")))
  richness[ ,i] <- rich
  
  # Calculate evenness
  even <- 
    numeric(
    matrix(estimate_richness(st.physeq.dna.r, measures = "InvSimpson")))
  evenness[ ,i] <- even
  
}
```

```{r}
#### Rarefied and filtered OTU table  ####
st.physeq.dna.rar # 425 taxa and 36 samples

sum(taxa_sums(st.physeq.dna.rar) == 0) #0
sum(taxa_sums(st.physeq.dna.rar) == 1) #28
sum(taxa_sums(st.physeq.dna.rar) == 2) #19
```

```{r}
# Reorder Type factor
TypeCategories = levels(get_variable(st.physeq.dna.rar, "Day"))
sample_data(st.physeq.dna.rar)$Day <- factor(sample_data(st.physeq.dna.rar)$Day, levels = c("0", "3", "7", "14", "21", "40"))

```

```{r}
# Create a new dataframe to hold the means and standard deviations of richness estimates
ID <- row.names(richness)
mean <- apply(richness, 1, mean)
sd <- apply(richness, 1, sd)
measure <- rep("Richness", nsamp)
rich_stats <- data.frame(ID, mean, sd, measure)
# Create a new dataframe to hold the means and standard deviations of evenness estimates
ID <- row.names(evenness)
mean <- apply(evenness, 1, mean)
sd <- apply(evenness, 1, sd)
measure <- rep("Inverse Simpson", nsamp)
even_stats <- data.frame(ID, mean, sd, measure)
# Combine our estimates for richness and evenness into one dataframe
alpha <- rbind(rich_stats, even_stats)
#write.table(alpha, file = "VF.rarefied.alpha.xls", sep = "\t", col.names=NA)
# Lets add the sample metadata into this dataframe using the merge() command
s <- data.frame(sample_data(st.physeq.dna.rar))
alphadiv <- merge(alpha, s, by = "ID") 
```

```{r}
# Plot 
ggplot(alphadiv, aes(x = Type, y = mean, color = Type, group = Type)) +
  geom_point(size = 2) + 
  geom_line(size = 0.8) #+
  #facet_wrap(Plant~Succession, scales = 'free_y') 
```

```{r
alpha_meas = c("Observed", "Chao1", "Shannon", "InvSimpson")
alpha_p <- plot_richness(Succ_cl.physeqFf.rar, "Plant", measures=alpha_meas)
alpha_p + geom_point(size = 4, alpha = 0.7)
alpha_p2 <- plot_richness(Succ_cl.physeqFf.rar, "Plant", color = "Succession", measures=alpha_meas)
alpha_p2 + geom_boxplot()

#richness - anova test
#load "a.diversity.anova.R"

#Plants only
Succ_cl.physeqFf.rar_noMG_Plants <- subset_samples(Succ_cl.physeqFf.rar_noMG, Type == "Plants")
rich_anova <- plot_anova_diversity(Succ_cl.physeqFf.rar_noMG_Plants, method = c("richness", "invsimpson"), 
                                  grouping_column = "Succession", pValueCutoff = 0.05)
p_alpha_anova <- rich_anova + scale_fill_viridis(discrete=TRUE, option = "E") + theme(axis.text.x=element_text(angle=0))


#DP - there are no "Succession" levels for Soil, one time point samples only
Succ_cl.physeqFf.rar_noMG_DP <- subset_samples(Succ_cl.physeqFf.rar_noMG, !Type == "Plants" & !Type == "Soil")
rich_anova_DP <- plot_anova_diversity(Succ_cl.physeqFf.rar_noMG_DP, method = c("richness", "invsimpson"), 
                                   grouping_column = "Succession", pValueCutoff = 0.1)
DP_alpha_anova <- rich_anova_DP + scale_fill_viridis(discrete=TRUE, option = "E") + theme(axis.text.x=element_text(angle=0))
```